Bridging [TBD]

### 2014-03-19 paper/draft.tex
Added draft of conclusion

% Integration NDSS Taming
Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor. The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution. DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach of demoting the greatest part of code to userspace. DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.

@@

Nested virtualization refers to a system architecture with two layers of virtualization. It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest. The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization.
%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.

Introducing nested virtualization, we have to identify the two main actors involved in the virtualization process: bare-metal hypervisor (L0) and nested hypervisor (L1). In addition, we will refer to the nested guest as L2 guest. 
However, virtualization introduces a computational overhead. It is small when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}. This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtles:ibm} proposing an architecture with acceptable overhead.

@@

\subsection{Conclusion}

@@

To sum up,  the adoption of MH could  increase feature set introduced by nested virtualization adoption. The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance. In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
 \section{Implementation}
In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction. In this section, we propose a GPH over MH implementation, based on Xen over NOVA.

@@

%Future works will firstly concern the study of available MH, trying to identify which project could be suitable to act as L0 in a GPH over MH platform. As underlined previously, the key feature of L0 MH will be the capacity of having a tiny TCB. Moreover, additional requirements concerns the flexibility in terms VMs ( number of VM supported at the same time, CPU handling, I/O management). Finally, another criteria for choosing the L0 hypervisor will be the evaluated effort needed for integrating a nested HV extension the to the base project.
%On the other hand, L1 choice is much simpler. Previous analysis identified Xen as the most suitable hypervisor for this role. The capacity to support different virtualization techniques added to the ability of running without HV support, makes Xen suitable to act as compatibility layer. 
%
%After this initial study, the first step will be deploying a minimal nested architecture testing ``out-of-the-box'' features and limitations. The following part will concern the conception of patch and extensions in order to adapt the general design of MH to properly act as L0.
%The final goal is to provide an architecture able to leverage nested virtualization cloud platforms improvements, having a smaller TCB and allowing customers to leverage flexible migrations and scheduling.

>>>>

% Integration NDSS Taming
Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor. The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution. DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach defining  the greatest part of code to userspace. DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.

@@

Nested virtualization refers to a system architecture with two layers of virtualization. It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest. The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization. Virtualization is able to execute in a deprivileged context  that normally requires a privileged execution. With nested virtualization we could apply this approach even to virtualization itself, executing a deprivileged hypervisor and leveraging code reutilization.
%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.

Introducing nested virtualization, we have to identify the two main actors involved in the virtualization process: bare-metal hypervisor (L0) and nested hypervisor (L1). In addition, we will refer to the nested guest as L2 guest. 
However, virtualization introduces a computational overhead. It is normally acceptable when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}. This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtles:ibm} proposing an architecture with acceptable overhead.

@@

\subsection{Lessons learned}

@@

To sum up,  the adoption of MH could  increase feature set introduced by nested virtualization adoption. The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance. In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
\section{Security Architecture}
\label{ref:arch}
\section{Vertical properties}

\section{Implementation}
In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction. In this section, we propose a GPH over MH implementation, based on Xen over NOVA \cite{nova}.

@@

In previous sections we provided a global analysis of modern challenges of cloud computing  and how those could be globally addressed addressed leveraging nested virtualization and different hypervisor architectures.

Future works will be focused on two complementary parts.
Firstly, they will concern the completion of the architecture of security in order to concretely meet horizontal design principles, identifying key components, similarly as done for the vertical properties.
At the same time, a major effort will be directed on enlarging the initial implementation.

%On the other hand, L1 choice is much simpler. Previous analysis identified Xen as the most suitable hypervisor for this role. The capacity to support different virtualization techniques added to the ability of running without HV support, makes Xen suitable to act as compatibility layer. 
%
%After this initial study, the first step will be deploying a minimal nested architecture testing ``out-of-the-box'' features and limitations. The following part will concern the conception of patch and extensions in order to adapt the general design of MH to properly act as L0.
%The final goal is to provide an architecture able to leverage nested virtualization cloud platforms improvements, having a smaller TCB and allowing customers to leverage flexible migrations and scheduling.
